Glue-on AdS holography for $T\bar T$-deformed CFTs

### 2023-06-29 alpha/GlueonAdS/glueon-poster.tex
Update poster [wip]

\documentclass[10pt]{article}

\input{preamble-slides.tex}

@@

\renewenvironment{frame}[1]%
	{\section*{#1}}%
	{\clearpage}

\newcommand{\pause}{}

\newenvironment{columns}%
	{\par\begin{minipage}{\textwidth}}%
	{\end{minipage}}

\newenvironment{column}[1]%
	{\begin{minipage}{#1}}%
	{\end{minipage}}

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
%\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}
\author{%
	Wen-Xin Lai
	\texstringonly{\,}\textkai{赖文昕}%
}
\date{}

@@

@@

\setlength{\droptitle}{-3.5\baselineskip}
\maketitle
\thispagestyle{empty}

\vspace{-3.5\baselineskip}

\centering

\textit{in collaboration with}
\\[.8ex]
	{\large Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}}
\\[2ex]
	\sidenote{\arxiv{2303.04836}}
\\[3ex]
	\textsf{Yau Mathematical Sciences Center, Tsinghua}
\\[1ex]
	{\large\ccbyncsajp}
\\[2ex]
	\textsf{\today\ @ BIMSA}
\\[4ex]
	\includegraphics[height=.38\textheight]{img/smbc-cantor-cropped.png}
\\[-.5ex]
	{\footnotesize\url{https://smbc-comics.com/comic/cantor}}

\clearpage
\justifying

%\raggedbottom
\large
\tableofcontents
\pagebreak

\section{\textbf{Warmup:} Pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$}

\newcommand{\citeMaldacena}{%
	\textcite{Maldacena:1997re}
}

\newcommand{\stateAdsCft}{
\begin{align*}
	\textrm{Strings on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{\textit{asympt.}~$\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at \textit{asympt.}~boundary}
\end{align*}
}

\citeMaldacena
\stateAdsCft
\begin{figure}[!h]
	\centering
	\includegraphics[width=.5\linewidth]{img/ads-cft.png}

	\vspace{-.2\baselineskip}
	\caption{$\mrm{AdS}_3/\mrm{CFT}_2$ cartoon}

	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{figure}
\vspace{-1.3\baselineskip}

%\begin{equation}
%	R_{\mu\nu} - \frac{1}{2}\,R g_{\mu\nu}
%	+ \Lambda g_{\mu\nu}
%	= 8\pi G_N T_{\mu\nu}^{\textit{matter}},\quad
%	\Lambda < 0
%\label{eq:einstein}
%\end{equation}

\section*{AdS/CFT --- a model of quantum gravity}
\citeMaldacena
\stateAdsCft
\begin{itemize}

\item A model of \textsc{quantum gravity}
\item ... originating from \textsc{string theory}
\item ... seemingly \textsc{Universal / ubiquitous}

\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}
	\item Our work: pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$ $(d=2)$.

\end{itemize}
\end{itemize}

\pagebreak

%%% LaTeX3e!
\NewDocumentCommand{\figAdsCft}{ O{.28} }{
\begin{minipage}{#1\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/ads-cft-cleaner.png}

	\vspace{-.01\baselineskip}
	\scriptsize\textcite{AldegundePWSep22}\ \,
\end{minipage}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a cylinder}

\textcite{Aharony:1999ti}

\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

	\vspace{-.5\baselineskip}
	boundary at $\rho \to 0$, $x^{i} \sim \varphi, t$, $i = 1,2$

\pause
	\item Surprisingly the expansion simply terminates at~$\rho\thinmspace g^{(4)}_{ij}$ for pure Einstein gravity! Moreover,
	\begin{equation}
		g^{(4)}_{ij} = \frac{1}{4} g^{(2)}_{ik} g^{(0)\thinmspace kl}_{\phantom{kl}} g^{(2)}_{lj}
	\end{equation}
	\textcite{Fefferman:2007rka,Banados:1998gg}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\newcommand{\eqBanados}{
	\begin{equation*}
	\hspace{-2em}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ \,
	\end{equation*}
}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Banados:1992wn}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric for $\rho > 0$, \\
	\textcite{Banados:1998gg}
	\eqBanados
	$g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
	where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\pause
	\item A consequence of Einstein equations \\
	in $(2+1)$ dimensions.\\
	\textit{Too many constraints, too little freedom!}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- simple realization of holography}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Different \textbf{geometries} in the ``bulk''
	\eqBanados
	specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: periodic functions, maps to different \textbf{states} in the boundary $\mrm{CFT}_2$ with stress energy $\ave{T(u)}_\psi$, $\ave{\bar{T}(v)}_\psi$

	\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities:\\
		\begin{itemize}
		\item spectrum,
			symmetry,\\
			partition functions,
			correlators, ...
		\end{itemize}
	\end{itemize}

\end{column}
\end{columns}
\end{frame}

\section{\textbf{Review:} \TTbar deformation and cutoff AdS} \label{se:cutoffholography}

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
\begin{columns}
\figAdsCft
\begin{column}{.72\textwidth}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\pause

@@

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-1.\baselineskip}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,
yet it's still mysterious and difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

\pause
\item What if we start from a $\mrm{CFT}_2$, that enjoys a holographic dual to $\mrm{AdS}_3$...

@@

\newcommand{\figCutoffAds}[1][.3]{
\begin{column}{#1\textwidth}
	\hfill\includegraphics[width=\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\flushleft
	\ \scriptsize\textcite{AldegundePWSep22}
\end{column}
\hspace{-2em}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-2em}
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item Question: what is the AdS dual of \TTbar deformation?\\

@@

\pause
	\item look at properties of \TTbar:\\
	signs of gravity and coarse-graining\\
	\textcite{Dubovsky:2012wk,Dubovsky:2013ira}
	\end{itemize}

	\pause
	\item Answer: AdS gravity within a finite Dirichlet wall

	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.7\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}
	\textit{Roughly $\mu \sim 1/r^2$}

\pause

	\item Passed many non-trivial tests
	\item Related to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\figCutoffAds
\hspace{-1.2em}
\begin{column}{.72\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

@@

\pause
	\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
	\end{equation}
	But \TTbar itself admits $\mu > 0$ with nice properties.\\
	\textit{What is the other side of the duality?}
		\begin{itemize}
		\item For comparison, the proposal of \textcite{Guica:2019nzm} admits both signs of $\mu$
		\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\newcommand{\figGlueon}[1][.37]{
\begin{minipage}{#1\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/diagram.pdf}
	\vspace{-1.5\baselineskip}
	\captionsetup{type=figure}
	\caption{Glue-on $\mrm{AdS}_3$}

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\figGlueon
\begin{column}{.6\textwidth}
\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

@@

\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\pause
\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together (``glue-on'');
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}\ \,
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}

\centering

\begin{minipage}{.5\textwidth}
\centering
\includegraphics[width=.9\linewidth]{img/diagram.pdf}
\end{minipage}

\begin{equation*}
\stateGlueon
\end{equation*}
\end{frame}

@@

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\figGlueon[.3]
\begin{column}{.7\textwidth}
\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\pause
\item Demanding the extended geometry to be non-singular
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\figGlueon[.3]
\begin{column}{.7\textwidth}
\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

\pause
\item We are careful to match the bulk \& boundary symmetries at $\mcal{N}_{\zeta_c}$

@@

\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
%\figGlueon[.3]
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Covariant charges of geometries
	\begin{align*}
		\delta E(\mu) &\equiv  \ell^{-1} \delta \mathcal Q_{\pd_{t'}}, \\ \delta J(\mu) &\equiv  \delta \mathcal Q_{\pd_{\varphi'}},
	\end{align*}

@@

\pause

\item This is realized by the \textit{state-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt,  \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar thermodynamics from bulk geometry}{%
	\textcite{Giveon:2017nie,Apolo:2019zai}
}
\begin{columns}
%\figGlueon[.3]
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}

\item \textit{State-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt, \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}

\pause
\item Many consequences: correct charges, signal propagation speed $v'_{\pm} \equiv \pm \frac{d\varphi'}{dt'}$, and thermodynamics when Wick rotated:
	\begin{equation}
		T_L(\mu)\,T_R(\mu) \le - \frac{1}{4\pi^2 \ell^2 \zeta_c} = \frac{3}{4\pi^2c\mu} = T_H(\mu)^2,
	\quad \mu > 0
	\end{equation}
	$T_{L,R}$ are temperatures associated with $u',v' = \varphi' \pm t'$.

\item $T_H$ is the \textit{Hagedorn temperature}: exceeding $T_H$ corresponds to a complex entropy
\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Caputa:2020lpa}
}
\begin{columns}
\figGlueon
\hspace{-1.5em}
\begin{column}{.6\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function is approximated by the on-shell gravity action of the dominant saddle:
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

\pause
\item $I(\zeta_c)$ is roughly the volume of the geometry, which diverges at $\zeta \to 0^\pm$. We apply a natural extension of \textit{holographic renormalization} to remove the divergence.

\pause
\item $I = -\log \mcal{Z}$ satisfies the \TTbar flow \eqref{TTbardef}.\\
	This is non-trivial because we've packaged the quantum corrections of the boundary theory.
\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\pause
\item We hope to understand the entanglement structure of \TTbar deformed theories from bulk geometry. 
\end{itemize}
\end{column}
\end{columns}
\end{frame}

>>>>

\documentclass[11pt]{article}

\input{preamble-poster.tex}

@@

\renewenvironment{frame}[1]%
	{\section*{#1}}%
	{}

%\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}

@@

\newcommand{\jokeInfinity}{
	\includegraphics[height=.2\linewidth]{img/smbc-cantor-cropped.png}
\\[-.5ex]
	{\footnotesize\url{https://smbc-comics.com/comic/cantor}}
}

\setlength{\droptitle}{-3.2\baselineskip}
\pretitle{\huge\bfseries\noindent}
\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\posttitle{\ \,}

\preauthor{\large}
\author{Wen-Xin Lai \textkai{赖文昕}}
\postauthor{, \textit{\footnotesize with}
	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}
	\ \arxiv{2303.04836}
	\par\vspace{.5\baselineskip}
	\hrule\vspace{.3\baselineskip}
}

\predate{\noindent\sffamily\,%
	Yau Mathematical Sciences Center YMSC, Tsinghua
	---
}
\date{July 2023}
\postdate{
	@ ICBS @ BIMSA
	\vspace{.5\baselineskip}
}

\pagestyle{empty}

@@

\maketitle
\thispagestyle{empty}

\begin{multicols}{3}

\section*{Foundation: pure Einstein in $\mrm{AdS}_3$~/~$\mrm{CFT}_2$}

\textcite{Maldacena:1997re}\\
\textcite{Aharony:1999ti}
\begin{align*}
	\textrm{Strings on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{\textit{asympt.}~$\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at \textit{asympt.}~boundary}
\end{align*}
\begin{center}
	\includegraphics[width=.6\linewidth]{img/ads-cft.png}

	\vspace{-.2\baselineskip}
	\captionsetup{type=figure}
	\caption{$\mrm{AdS}_3/\mrm{CFT}_2$ cartoon}

	\vspace{.1\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{center}

\item \textsc{Universal / ubiquitous}

\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}

\item Our work: pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$ $(d=2)$.\\
$\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
\begin{equation}
\hspace{-2em}
\begin{aligned}
	ds^2
	&= \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j
\\[.5ex]
	&= \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)
	\label{fggauge}
\end{aligned}
\end{equation}

\vspace{-.5\baselineskip}
boundary at $\rho \to 0$, $x^{i} \sim \varphi, t$, indices $i,j = 1,2$

\item Surprisingly the expansion simply terminates at~$\rho\thinmspace g^{(4)}_{ij}$ for pure Einstein gravity! Moreover,
\begin{equation}
	g^{(4)}_{ij} = \frac{1}{4} g^{(2)}_{ik} g^{(0)\thinmspace kl}_{\phantom{kl}} g^{(2)}_{lj}
\end{equation}
\textcite{Fefferman:2007rka}\\
\textcite{Banados:1992wn}\\
\textcite{Banados:1998gg}
\end{itemize}

\newcommand{\eqBanados}{
	\begin{equation*}
	\hspace{-2em}
		ds^2 \ \,
	\end{equation*}
}

\begin{itemize}
\item $g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\item A consequence of Einstein equations \\
in $(2+1)$ dimensions.\\
\textit{Too many constraints, too little freedom!}
\end{itemize}

\begin{itemize}
\item Different \textbf{geometries} in the ``bulk''

specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: periodic functions, maps to different \textbf{states} in the boundary $\mrm{CFT}_2$ with stress energy $\ave{T(u)}_\psi$, $\ave{\bar{T}(v)}_\psi$

\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities:
	\begin{itemize}
	\item spectrum,
		symmetry,\\
		partition functions,
		correlators, ...
	\end{itemize}
\end{itemize}

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{frame}

@@

@@

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-.1\baselineskip}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,
yet it's still mysterious and difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

\item What if we start from a $\mrm{CFT}_2$, that enjoys a holographic dual to $\mrm{AdS}_3$...

@@

\newcommand{\figCutoffAds}[1][.5]{
\begin{center}
	\hspace{.15\linewidth}
	\includegraphics[width=#1\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\ \scriptsize\textcite{AldegundePWSep22}
\end{center}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}

\figCutoffAds

\begin{itemize}
\item Question: what is the AdS dual of \TTbar deformation?

@@

\item look at properties of \TTbar:\\
signs of gravity and coarse-graining\\
\textcite{Dubovsky:2012wk,Dubovsky:2013ira}
\end{itemize}

\item Answer: AdS gravity within a finite Dirichlet wall

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{}

	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}
	\textit{Roughly $\mu \sim 1/r^2$}

	\item Passed many non-trivial tests,\\
	related to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}

\begin{itemize}
\item \textbf{Significance:} gravity inside a finite box\\
	with hard Dirichlet walls

@@

\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
\begin{equation}
	\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
\end{equation}
But \TTbar itself admits $\mu > 0$ with nice properties.\\
\textit{What is the other side of the duality?}
	\begin{itemize}
	\item For comparison, the proposal of \textcite{Guica:2019nzm} admits both signs of $\mu$
	\end{itemize}
\end{itemize}

@@

\newcommand{\figGlueon}[1][1]{
\begin{minipage}{#1\linewidth}
	\centering
	\includegraphics[width=.5\linewidth]{img/diagram.pdf}
	\vspace{-1.5\baselineskip}
	\captionsetup{type=figure}
	\caption{Glue-on $\mrm{AdS}_3$}

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}

\figGlueon

\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

@@

@@

\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together (``glue-on'');
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{}

\begin{center}
\includegraphics[width=.6\linewidth]{img/diagram.pdf}
\begin{equation*}
\stateGlueon
\end{equation*}
\end{center}

@@

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}

\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\item Demanding the extended geometry to be non-singular
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}

\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

\item We are careful to match the bulk \& boundary symmetries at $\mcal{N}_{\zeta_c}$

@@

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}

\begin{itemize}
\item Covariant charges of geometries
	\begin{align*}
		\delta E(\mu) &\equiv  \ell^{-1} \delta \mathcal Q_{\pd_{t'}}, \\ \delta J(\mu) &\equiv  \delta \mathcal Q_{\pd_{\varphi'}},
	\end{align*}

@@

\item This is realized by the \textit{state-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt,  \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}
\end{itemize}

\begin{frame}{\TTbar thermodynamics from bulk geometry}{%
	\textcite{Giveon:2017nie,Apolo:2019zai}
}

\item Many consequences: correct charges, signal propagation speed $v'_{\pm} \equiv \pm \frac{d\varphi'}{dt'}$, and thermodynamics when Wick rotated:
	\begin{equation}
		T_L(\mu)\,T_R(\mu) \le - \frac{1}{4\pi^2 \ell^2 \zeta_c} = \frac{3}{4\pi^2c\mu} = T_H(\mu)^2,
	\quad \mu > 0
	\end{equation}
	$T_{L,R}$ are temperatures associated with $u',v' = \varphi' \pm t'$.

\item $T_H$ is the \textit{Hagedorn temperature}: exceeding $T_H$ corresponds to a complex entropy
\end{itemize}

@@

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Caputa:2020lpa}
}

\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function is approximated by the on-shell gravity action of the dominant saddle:
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

\item $I(\zeta_c)$ is roughly the volume of the geometry, which diverges at $\zeta \to 0^\pm$. We apply a natural extension of \textit{holographic renormalization} to remove the divergence.

\item $I = -\log \mcal{Z}$ satisfies the \TTbar flow \eqref{TTbardef}.\\
	This is non-trivial because we've packaged the quantum corrections of the boundary theory.
\end{itemize}
\end{frame}

@@

\item We hope to understand the entanglement structure of \TTbar deformed theories from bulk geometry. 
\end{itemize}
